### 2021-02-04 manuscript/appendix.Rmd
initial commit of manuscript

# Appendix A: Protocol Changes {#appA}

# Appendix B title {#appB}

Tabular data for citation curves and citation valence

```{r results = "asis"}
options(knitr.kable.NA = "-") # set symbol to use for NA cells

d_summary %>%
  filter(case == 'baumeister' & pubYear %in% c(2015,2017,2018,2019) |
         case == 'strack' & pubYear %in% c(2015,2017,2018,2019) |
         case == 'sripada' & pubYear %in% c(2015,2017,2018,2019) |
         case == 'carter' & pubYear %in% c(2013,2015,2016,2017,2018,2019) |
         case == 'caruso' & pubYear %in% c(2013,2015,2016,2017,2018,2019)) %>%
  mutate(favourable = paste0(favourable,' (',round(favourable_prop*100,0),'%)'),
         unfavourable = paste0(unfavourable,' (',round(unfavourable_prop*100,0),'%)'),
         equivocal = paste0(equivocal,' (',round(equivocal_prop*100,0),'%)'),
         unclassifiable = paste0(unclassifiable,' (',round(unclassifiable_prop*100,0),'%)'),
         excluded = paste0(excluded_yes,' (',round(excluded_yes_prop*100,0),'%)'),
         citesRep = paste0(citesRep_yes,' (',round(citesRep_yes_prop*100,0),'%)'),
         citesRep = ifelse(case %in% c('baumeister','sripada','strack') & pubYear == 2015, NA, citesRep), # set citesRep to NA in pre-replication year
         citesRep = ifelse(case %in% c('carter','caruso') & pubYear == 2013, NA, citesRep)) %>% # set citesRep to NA in pre-replication year
  select(year = pubYear, `citing\narticles` = citesTarget, excluded, unclassifiable, favourable, equivocal, unfavourable, citesRep) %>%
  kable(caption = 'Counts (percentages) for citations to the original study, exclusions, citation valence, and concomitant citations to replication study in the time periods subjected to qualitative analysis. Percentages for the Baumeister case are extrapolations based on 40\\% random samples drawn from the pre-replication and post-replication time periods.', booktabs=T, longtable=T) %>%
  pack_rows("Baumeister", 1, 4) %>%
  pack_rows("Sripada", 5, 8) %>%
  pack_rows("Strack", 9, 12) %>%
  pack_rows("Carter", 13, 18) %>%
  pack_rows("Caruso", 19, 24)
```

```{r}
# DATA SUMMARISED FOR POST-REPLICATION PERIOD
d_summary %>%
  filter(year == 'post-replication') %>%
  mutate(favourable = paste0(favourable,' (',round(favourable_prop,2),')'),
         unfavourable = paste0(unfavourable,' (',round(unfavourable_prop,2),')'),
         equivocal = paste0(equivocal,' (',round(equivocal_prop,2),')'),
         unclassifiable = paste0(unclassifiable,' (',round(unclassifiable_prop,2),')'),
         excluded = paste0(excluded_yes,' (',round(excluded_yes_prop,2),')'),
         citesRep = paste0(citesRep_yes,' (',round(citesRep_yes_prop,2),')')) %>%
  select(case, year = pubYear, `citing\narticles` = citesTarget, excluded, unclassifiable, favourable, equivocal, unfavourable, citesRep)
```